What's in a name? On the (automated) topic naming of software maintenance activities

### 2010-02-24 icpc/icpc2010.tex
minor icp2010 changes to get it to compile
just figuring out what to do

\newtheorem{mydef}{definition}
\usepackage{graphicx}
\newcommand{\slbl}[1]{\textbf{\textsf{#1}}}
\usepackage{multirow}%,multicolumn}
\usepackage{slashbox}
\usepackage{wrapfig}
\usepackage{booktabs, colortbl}
\usepackage{amsmath,amssymb}
\usepackage{algpseudocode}
\usepackage[pdfstartview=FitH,xetex,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
\usepackage{algorithm}%,algorithmic-fix}
\usepackage{url}
\begin{document}

@@

\begin{abstract}
  In automated mining of software repositories, most approaches, such as topic modeling and concept location, focus on applying standard machine learning algorithms to corpora. This approach does not make any use of domain-specific information to better understand results. While too much specificity can produce non-generalizable results, too little produces broad learners that don't provide much useful detail.%There has not been any   investigation into the latent topics these classifiers represent. 
In this paper, we investigate the latent topics classifier produce, and examine  whether such topics can be
  generalized across software products. We focus on software quality
  as a potential generalization, since there is some shared belief
  that these qualities apply broadly across software products. We
  wanted to name topics based on a static ontology of non-functional
  requirements and software qualities. %Our overarching goal is to use   individual topic models to create a list of terms that can be shared   across software projects. 
In essence, this would be adding some form
  of context to general-purpose ontologies such as Wordnet. In this
  paper, we discuss how to use domain knowledge about software
  development in order to better annotate topics of development that
  are extracted from the change log messages in configuration
  management systems such as CVS. Since our taxonomy is global, our results allow us to compare the relative importance of particular qualities between projects.
\end{abstract}

@@

Unlike the poetic licence Shakespeare uses, we show that names are indeed important. In this paper we `name' topic models generated with a topic modeling technique, Latent Dirichlet Allocation~\cite{Blei2003}. We use these topic models to train supervised learners to classify larger datasets. We also show that these topics can be generalized across products.

Traditionally topic modeling has required manual annotation to derive domain-relevant information. This paper introduces \emph{named topic models}, topic models which are named using a cross-project taxonomy. We start with an unsupervised topic learning algorithm, then manually annotate the topics using the ontology. From the annotated topics, we apply automated classifiers to find labels for the remaining topics.

@@

Our contributions include:
\begin{itemize}
\item  We introduce named topic analysis, and demonstrate its usefulness
  compared to other approaches.
\item We show that these topics can be learned and used to classify other datasets.
\item We present a number of visualizations of named topics and their trends over time to aid
  communication and analysis.
\item We use an exploratory case study of
  several open source database systems to show how named topics can be compared between projects.
\end{itemize}

@@

\section{Background}
We provide a brief overview of software repository mining and information retrieval.
\subsection{Definitions}
A \emph{message} is a block of text written by a developer. In this
paper, messages will be the CVS and BitKeeper commit log comments made
when the user commits changes to files in a repository. A \emph{word
  distribution} is the summary of a message by word count. Each word
distribution will be over the words in all messages. However, most
words will not appear in each message. A word distribution is effectively
a word count divided by the message size. A \emph{topic} is a word
distribution, i.e., a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top 10 most frequent words of their word distribution.  A \emph{trend}
is one or more similar topics that recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project.

\emph{ROC} values reflect a score, similar to school letter-grades, reflecting how well a particular learner performed for the given data. A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data). ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) vs. the false positive rate (1 - specificity). A perfect learner has a value of 1.0, reflecting perfect recall and precision.

@@

To construct our expanded sets, we also used Wordnet~\cite{Fellbaum1998}, an English-language ‘lexical database’ that contains semantic relations between words, including meronymy and synonymy. We then added Boehm’s 1976 software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven ‘ilities’ into their respective ISO9126 qualities. We did the same for the quality model produced by McCall et al. \cite{mccall1977}. Finally, we analyzed two mailing lists from the KDE project to enhance the specificity of the sets. We selected KDE-Usability, which focuses on usability discussions for KDE as a whole; and KDE-Konqueror, a list about a long-lived web browser project. For each high-level quality in ISO9126, we first searched for our existing labels; we then randomly sampled twenty-five mail messages that were relevant to that quality, and selected co-occurring terms relevant to that quality. For example, we add the term ‘performance’ to the synonyms for \emph{Efficiency}, since this term occurs in most KDE mail messages that discuss efficiency.
%exp2

@@

To assess the performance of the supervised learners, we did a 20-fold cross-over analysis (using 1 period to train the learner on the tags, and the other 19 to attempt to learn, changing the training set each time). We report these results below.

Finally, using this data, we evaluated two research questions: 1) Do label frequencies change over time? That is, is a certain quality of more interest at one point in the lifecycle than some other? 2) Do the different projects differ in their relative topic interest? That is, is a particular quality more important to one project than the other projects?

@@

\subsection{Word list similarity}
In general, this approach did not work out well. The related words for \emph{correctness}, for example, tended to be too lengthy and non-specific. Table \ref{tbl:wordlist} lists results. A \emph{named topic} is a topic with a matching label. There are \{periods X 20\} topics per project. All experiments were run on MaxDB data.

>>>>

\newtheorem{mydef}{definition}
\usepackage{graphicx}
\newcommand{\slbl}[1]{\textbf{\textsf{#1}}}
\usepackage{multirow}%,multicolumn}
\usepackage{slashbox}
\usepackage{wrapfig}
\usepackage{booktabs, colortbl}
\usepackage{amsmath}
%\usepackage{amsmath,amssymb}
%\usepackage{algpseudocode}
\usepackage[pdfstartview=FitH,xetex,bookmarksopenlevel=3,bookmarks=true]{hyperref} %pdfpagescrop={92 112 523 778},
%\usepackage{algorithm}%,algorithmic-fix}
\usepackage{url}
\begin{document}

@@

\begin{abstract}
  In the automated mining of software repositories, most approaches, such as topic modeling and concept location, focus on applying standard machine learning algorithms to corpora. This approach does not make any use of domain-specific information to better understand results. While too much specificity can produce non-generalizable results, too little produces broad learners that don't provide much useful detail.%There has not been any   investigation into the latent topics these classifiers represent. 
In this paper, we investigate what topics the latent topics classifier produce, and examine  whether such topics can be
  generalized across software products. We focus on nonfunctional requirements related to software quality
  as a potential generalization, since there is some shared belief
  that these qualities apply broadly across software products. We
  wanted to name topics based on a static ontology of non-functional
  requirements and software qualities. %Our overarching goal is to use   individual topic models to create a list of terms that can be shared   across software projects. 
In essence, this would be adding some form
  of context to general-purpose ontologies such as Wordnet. In this
  paper, we discuss how to use domain knowledge about software
  development in order to better annotate topics of development that
  are extracted from the change log messages in configuration
  management systems such as CVS. Since our taxonomy is global, our results allow us to compare the relative importance of particular qualities between projects.
\end{abstract}

@@

Unlike the poetic licence Shakespeare uses, we show that names are indeed important. In this paper we `name' and label topic models generated by a topic modeling technique, Latent Dirichlet Allocation~\cite{Blei2003}. We use these topic models to train supervised learners to classify larger datasets. We also show that these topics can be generalized across products.

% Clean this up
Traditionally topic modeling has required manual annotation to derive domain-relevant information. This paper introduces \emph{named topic models}, topic models which are named using a cross-project taxonomy. We start with an unsupervised topic learning algorithm, then manually annotate the topics using the ontology. From the annotated topics, we apply automated classifiers to find labels for the remaining topics.

@@

Our contributions include:
\begin{itemize}
\item  We introduce named topic analysis, and demonstrate its usefulness
  compared to other approaches.
% unsure about this
\item We show that these topics can be learned and used to classify other datasets.
\item We present a number of visualizations of named topics and their trends over time to aid
  communication and analysis.
\item We use an exploratory case study of
  several open source database systems to show how named topics can be compared between projects.
\end{itemize}

@@

\section{Background}

We provide a brief overview of software repository mining and information retrieval.

\subsection{Definitions}
A \emph{message} is a block of text written by a developer. In this
paper, messages will be the CVS and BitKeeper commit log comments made
when the user commits changes to files in a repository. A \emph{word
  distribution} is the summary of a message by word count. Each word
distribution will be over the words in all messages. However, most
words will not appear in each message. A word distribution is effectively
a word count divided by the message size. A \emph{topic} is a word
distribution, i.e., a set of words that form a word distribution that is
unique and independent within the set of documents in our total
corpus. One could think of a topic as a distribution of the centroid
of a group of messages. In this paper we often summarize topics by the
top 10 most frequent words of their word distribution.  A \emph{trend}
is one or more similar topics that recur over time.  Trends are
particularly important, as they indicate long-lived and recurring
topics that may provide key insights into the development of the
project.

% area under the ROC 
% check
\emph{Area of ROC Curve} is the area under the Reciever Operating Characteric curve, we will refer to this as \emph{ROC}. ROC values reflect a score, similar to school letter-grades, reflecting how well a particular learner performed for the given data. A ROC result of 0.5 would be equivalent to a random learner (randomly classifying data). ROC maps to the more familiar concepts of precision/sensitivity and recall/specificity: it plots the true positive rate (sensitivity) vs. the false positive rate (1 - specificity). A perfect learner has a value of 1.0, reflecting perfect recall and precision.

@@

%Cite mockus et al
To construct our expanded sets, we also used Wordnet~\cite{Fellbaum1998}, an English-language ‘lexical database’ that contains semantic relations between words, including meronymy and synonymy. We then added Boehm’s 1976 software quality model \cite{Boehm+:1976:ICSE}, and classified his eleven ‘ilities’ into their respective ISO9126 qualities. We did the same for the quality model produced by McCall et al. \cite{mccall1977}. Finally, we analyzed two mailing lists from the KDE project to enhance the specificity of the sets. We selected KDE-Usability, which focuses on usability discussions for KDE as a whole; and KDE-Konqueror, a list about a long-lived web browser project. For each high-level quality in ISO9126, we first searched for our existing labels; we then randomly sampled twenty-five mail messages that were relevant to that quality, and selected co-occurring terms relevant to that quality. For example, we add the term ‘performance’ to the synonyms for \emph{Efficiency}, since this term occurs in most KDE mail messages that discuss efficiency.
%exp2

@@

To assess the performance of the supervised learners, we did a 10-fold cross-over analysis (using 1 period to train the learner on the tags, and the other 9 to attempt to learn, changing the training set each time). We report these results below.

% enumerate?
Finally, using this data, we evaluated two research questions: 1) Do label frequencies change over time? That is, is a certain quality of more interest at one point in the lifecycle than some other? 2) Do the different projects differ in their relative topic interest? That is, is a particular quality more important to one project than the other projects?

@@

\subsection{Word list similarity}

% word list similarity

In general, this approach did not work out well. The related words for \emph{correctness}, for example, tended to be too lengthy and non-specific. Table \ref{tbl:wordlist} lists results. A \emph{named topic} is a topic with a matching label. There are \{periods X 20\} topics per project. All experiments were run on MaxDB data.
